Weighted Model Counting/Integration from the Perspective of Boolean Algebras

### 2020-03-20 paper.tex
A nice diagram

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{conjecture}{Conjecture}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

@@

\DeclareMathOperator{\WMC}{WMC}
\DeclareMathOperator{\nWMC}{nWMC}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\End}{End}

@@

\subsection{Preliminaries}

\begin{definition} \label{def:ba}
  A \emph{Boolean algebra} (BA) is a tuple $(\mathbf{B}, \land, \lor, \neg, 0,
  1)$ of a set $\mathbf{B}$ with operations $\land, \lor, \neg$ and elements $0,
  1 \in \mathbf{B}$ such that the following axioms hold for all $a, b, \in
  \mathbf{B}$:
  \begin{itemize}
  \item both $\land$ and $\lor$ are associative and commutative;
  \item $a \lor (a \land b) = a$, and $a \land (a \lor b) = a$;
  \item $0$ is the identity of $\lor$, and $1$ is the identity of $\land$;
  \item $\lor$ distributes over $\land$ and vice versa;
  \item $a \lor \neg a = 1$, and $a \land \neg a = 0$.
  \end{itemize}

  We can also define a partial order $\le$ on $\mathbf{B}$ as $a \le b$ if $a =
  b \land a$ (or, equivalently, $a \lor b = b$) for $a, b \in \mathbf{B}$.
  Furthermore, let $a < b$ denote $a \le b$ and $a \ne b$.
\end{definition}

For the rest of this paper, let $\mathbf{B}$ refer to the BA $(\mathbf{B},
\land, \lor, \neg, 0, 1)$.

@@

@@

% TODO: all finite Boolean algebras are isomorphic to fields of sets

% TODO: we don't have to treat 0 as an exception we state that \/ of an empty
% set is 0 (and /\ of an empty set is 1).
% TODO: use bigvee and bigwedge directly with sets.
\begin{theorem} \label{thm:representation}
  Let $\mathbf{B}$ be a finite Boolean algebra. Then every $x \in \mathbf{B}
  \setminus \{ 0 \}$ can be uniquely expressed as
  \[
  x = \bigvee_{\text{atoms } a \le x} a.
  \]
\end{theorem}
\begin{proof}
  A simple consequence of the theorem that every finite Boolean algebra is
  isomorphic to a field of subsets of a set, where the cardinality of the set is
  equal to the number of atoms in the Boolean algebra.
\end{proof}

\todo[inline]{Remove the requirement for being strictly positive}
\begin{definition}[\cite{gaifman1964concerning}] \label{def:measure}
  A \emph{(strictly positive) measure} on a Boolean algebra $\mathbf{B}$ is a
  function $m\colon \mathbf{B} \to [0, 1]$ such that:
  \begin{enumerate}
  \item $m(1) = 1$, and $m(x) > 0$ for $x \ne
    0$; \label[property]{property:values}
  \item $m(x \lor y) = m(x) + m(y)$ for all $x, y \in \mathbf{B}$ whenever $x
    \land y = 0$. \label[property]{property:additive}
  \end{enumerate}
\end{definition}

% TODO: Feedback: Probably a good idea to explain how elements of the Boolean
% algebra map to models in WMC formulations. Probably do an example:
% 0.3 p; 0.2 q => M = <p, qbar> has weight o3 * (1-0.2)
% then it corresponds to an atom in an algebra defined over the language defined
% over propositions p and q.
% TODO:  Mention Lindenbaum-Tarski algebra somewhere

\subsection{New Results}

\todo[inline]{Allow weight to be zero}
\begin{definition} \label{def:wmc}
  Let $\mathbf{B}$ be a finite Boolean algebra, and let $M \subseteq \mathbf{B}$
  be its set of atoms. Let $L \subseteq \mathbf{B}$ be such that every atom $m
  \in M$ can be uniquely expressed as $m = \bigwedge_{i \in I} l_i$ for some $\{
  l_i \}_{i \in I} \subseteq L$, and let $w\colon L \to \mathbb{R}_{>0}$ be
  arbitrary. The \emph{weighted model count} $\WMC\colon \mathbf{B} \to
  \mathbb{R}_{\ge 0}$ is defined as
  \[
    \WMC(a) = \begin{cases}
      0 & \text{if } a = 0 \\
      \prod_{i \in I} w(l_i) & \text{if } M \ni a = \bigwedge_{i \in I} l_i \\
      \sum_{i \in I} \WMC(m_i) & \text{if } \mathbf{B} \setminus (M \cup \{ 0
      \}) \ni a = \bigvee_{i \in I} m_i
    \end{cases}
  \]
  for any $a \in \mathbf{B}$. Furthermore, we define the \emph{normalised
    weighted model count} $\nWMC\colon \mathbf{B} \to [0, 1]$ as $\nWMC(a) =
  \frac{\WMC(a)}{\WMC(1)}$ for all $a \in \mathbf{B}$.
\end{definition}
% TODO: Feedback: maybe NWMC or WMC^*, as nWMC looks like a typo. Also clarify
% the relationship between NWMC and the probability induced by a WMC
% formulation, i.e., Pr_{\Delta, w}(q) = \frac{WMC(\Delta \land q,
% w)}{WMC(\Delta, w)}.

\begin{proposition}
  $\nWMC$ is a measure for any finite Boolean algebra $\mathbf{B}$.
\end{proposition}
\begin{proof}
  First, note that \cref{property:values} of \cref{def:measure} is satisfied by
  the definition of $\nWMC$. Next, in order to prove \cref{property:additive},
  let $x, y \in \mathbf{B}$ be such that $x \land y = 0$. We want to show that
  \[
    \nWMC(x \lor y) = \nWMC(x) + \nWMC(y)
  \]
  which is equivalent to
  \begin{equation} \label{eq:additivity_proof}
    \WMC(x \lor y) = \WMC(x) + \WMC(y).
  \end{equation}
  If, say, $x = 0$, then \cref{eq:additivity_proof} becomes
  \[
    \WMC(y) = \WMC(0) + \WMC(y) = \WMC(y)
  \]
  (and likewise for $y = 0$). Thus we can assume that $x \ne 0 \ne y$ and use
  \cref{thm:representation} to write
  \[
    x = \bigvee_{i \in I} x_i \quad \text{and} \quad y = \bigvee_{j \in J} y_j
  \]
  for some sequences of atoms $(x_i)_{i \in I}$ and $(y_j)_{j \in J}$. If
  $x_{i'} = y_{j'}$ for some $i' \in I$ and $j' \in J$, then
  \[
    x \land y = \bigvee_{i \in I} \bigvee_{j \in J} x_i \land y_j = x_{i'} \land
    y_{j'} \ne 0,
  \]
  contradicting the assumption. This is enough to show that
  \begin{align*}
    \WMC(x \lor y) &= \WMC\left( \left( \bigvee_{i \in I} x_i \right) \lor \left(\bigvee_{j \in J} y_j \right) \right) = \sum_{i \in I} \WMC(x_i) + \sum_{j \in J} \WMC(y_j) \\
                   &= \WMC(x) + \WMC(y),
  \end{align*}
  finishing the proof.
\end{proof}

@@

\begin{proposition}
  Let $\mathbf{B}$ be a finite measure algebra with measure $m\colon \mathbf{B} \to
  \mathbb{R}_{\ge 0}$. Let $L \subset \mathbf{B}$ be defined as
  \[
  L = \{ l_i \mid i \in [n] \} \cup \{ \neg l_i \mid i \in [n] \}
  \]
  for some $n \in \mathbb{N}$. Finally, assume that $\mathbf{B}$ has $2^n$
  atoms, where each atom $a \in \mathbf{B}$ is an infimum
  \[
    a = \bigwedge_{i=1}^n a_i
  \]
  such that $a_i \in \{ l_i, \neg l_i \}$ for $i \in [n]$. Then there exists a
  weight function $w\colon L \to \mathbb{R}_{>0}$ that makes $m$ a WMC measure if
  and only if
  \begin{equation} \label{eq:wmccondition}
  m(l \land l') = m(l)m(l')
  \end{equation}
  for all distinct $l, l' \in L$ such that $l \ne \neg l'$.
\end{proposition}

@@

\begin{proof}
  Let us begin with the `if' part of the statement. Let $w\colon L \to
  \mathbb{R}_{>0}$ be defined by
  \begin{equation} \label{eq:assumption}
    w(l) = m(l)
  \end{equation}
  for all $l \in L$. We are going
  to show that $\nWMC = m$. First, note that $\nWMC(0) = 0 = m(0)$ by the
  definitions of both $\nWMC$ and $m$. Second, let
  \begin{equation} \label{eq:def_of_a}
    a = \bigwedge_{i=1}^n a_i
  \end{equation}
  be an atom in $\mathbf{B}$ such that $a_i \in \{ l_i, \neg l_i \}$ for all $i
  \in [n]$. Then
  \[
    \nWMC(a) = \frac{\WMC(a)}{\WMC(1)} = \frac{1}{\WMC(1)} \prod_{i=1}^n w(a_i)
    = \frac{1}{\WMC(1)} \prod_{i=1}^n m(a_i) = \frac{1}{\WMC(1)} m \left(
      \bigwedge_{i=1}^n a_i \right) = \frac{m(a)}{\WMC(1)}
  \]
  by \cref{def:wmc,eq:assumption,eq:wmccondition,eq:def_of_a}. Now we just need
  to show that $\WMC(1) = 1$. Indeed,
  \begin{align*}
    \WMC(1) &= \sum_{\text{atoms } a \in \mathbf{B}} \WMC(a) = \sum_{\text{atoms
      } a \in \mathbf{B}} \prod_{i=1}^n w(a_i) = \sum_{\text{atoms } a \in
      \mathbf{B}} \prod_{i=1}^n m(a_i) \\
    &= \sum_{\text{atoms } a \in
      \mathbf{B}} m \left( \bigwedge_{i=1}^n a_i \right) = \sum_{\text{atoms } a
      \in \mathbf{B}} m(a) = m \left( \bigvee_{\text{atoms } a \in \mathbf{B}}
    \right) = m(1) = 1.
  \end{align*}
  Finally, note that if $\nWMC$ and $m$ agree on all atoms, then they must also
  agree on all other non-zero elements of the Boolean algebra.

  For the other direction, we are given a weight function $w\colon L \to
  \mathbb{R}_{>0}$ that induces a measure $m = \nWMC\colon \mathbf{B} \to
  \mathbb{R}_{\ge 0}$, and we want to show that \cref{eq:wmccondition} is
  satisfied. Let $k_i, k_j \in L$ be such that $k_i \in \{ l_i, \neg l_i \}$,
  $k_j \in \{ l_j, \neg l_j \}$, and $i \ne j$. We will first prove an auxiliary
  result that
  \begin{equation} \label{eq:to_prove}
    m(k_i \land k_j) = m(k_i)m(k_j)
  \end{equation}
  is equivalent to
  \begin{equation} \label{eq:to_prove2}
    m(k_i \land k_j) \cdot m(\neg k_i \land \neg k_j) = m(k_i \land \neg k_j)
    \cdot m(\neg k_i \land k_j).
  \end{equation}
  First, note that $k_i$ can be expressed as
  \[
    k_i = (k_i \land k_j) \lor (k_i \land \neg k_j)
  \]
  with the condition that
  \[
    (k_i \land k_j) \land (k_i \land \neg k_j) = 0,
  \]
  so, by properties of a measure,
  \begin{equation} \label{eq:temp}
    m(k_i) = m(k_i \land k_j) + m(k_i \land \neg k_j).
  \end{equation}
  Applying \cref{eq:temp} and the equivalent expression for $m(k_j)$ allows us
  to rewrite \cref{eq:to_prove} as
  \begin{align*}
    m(k_i \land k_j) &= [m(k_i \land k_j) + m(k_i \land \neg k_j)] \cdot [m(k_i \land k_j) + m(\neg k_i \land k_j)] \\
                     &= m(k_i \land k_j)^2 + m(k_i \land k_j)[m(k_i \land \neg k_j) + m(\neg k_i \land k_j)] + m(k_i \land \neg k_j)m(\neg k_i \land k_j)
  \end{align*}
  Dividing both sides by $m(k_i \land k_j)$ gives
  \begin{equation} \label{eq:temp2}
    1 = m(k_i \land k_j) + m(k_i \land \neg k_j) + m(\neg k_i \land k_j) +
    \frac{m(k_i \land \neg k_j)m(\neg k_i \land k_j)}{m(k_i \land k_j)}.
  \end{equation}
  Since $k_i \land k_j \land k_i \land \neg k_j = 0$, and
  \[
    (k_i \land k_j) \lor (k_i \land \neg k_j) = k_i \land (k_j \lor \neg k_j) =
    k_i \land 1 = k_i,
  \]
  we have that
  \[
    m(k_i \land k_j) + m(k_i \land \neg k_j) = m(k_i).
  \]
  Similarly, $k_i \land \neg k_i \land k_j = 0$, and
  \[
    k_i \lor (\neg k_i \land k_j) = (k_i \lor \neg k_i) \land (k_i \lor k_j) =
    k_i \lor k_j,
  \]
  so
  \[
    m(k_i) + m(\neg k_i \land k_j) = m(k_i \lor k_j).
  \]
  Finally, note that
  \[
    (k_i \lor k_j) \land \neg(k_i \lor k_j) = 0,
  \]
  and
  \[
    (k_i \lor k_j) \lor \neg(k_i \lor k_j) = 1,
  \]
  so
  \[
    m(k_i \lor k_j) + m(\neg(k_i \lor k_j)) = m(1) = 1.
  \]
  This allows us to rewrite \cref{eq:temp2} as
  \[
    \frac{m(k_i \land \neg k_j)m(\neg k_i \land k_j)}{m(k_i \land k_j)} = 1 -
    m(k_i \lor k_j) = m(\neg(k_i \lor k_j)) = m(\neg k_i \land \neg k_j)
  \]
  which immediately gives us \cref{eq:to_prove2}.

@@

A well-known way to overcome this limitation of independence is by adding more
literals \cite{DBLP:journals/ai/ChaviraD08}, i.e., extending the set $L$ covered
by the WMC weight function $w\colon L \to \mathbb{R}_{>0}$. Let us translate this
idea to the language of Boolean algebras.

>>>>

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{conjecture}{Conjecture}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\definecolor{color1}{HTML}{fbb4ae}
\definecolor{color2}{HTML}{b3cde3}
\definecolor{color3}{HTML}{ccebc5}
\definecolor{color4}{HTML}{decbe4}

@@

\DeclareMathOperator{\WMC}{WMC}
\DeclareMathOperator{\nWMC}{NWMC}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\End}{End}

@@

\section{Preliminaries}

\begin{definition} \label{def:ba}
  A \emph{Boolean algebra} (BA) is a tuple $(\mathbf{B}, \land, \lor, \neg, 0,
  1)$ consisting of a set $\mathbf{B}$ with binary operations \emph{meet}
  $\land$ and \emph{join} $\lor$, unary operation $\neg$ and elements $0, 1 \in
  \mathbf{B}$ such that the following axioms hold for all $a, b, \in
  \mathbf{B}$:
  \begin{itemize}
  \item both $\land$ and $\lor$ are associative and commutative;
  \item $a \lor (a \land b) = a$, and $a \land (a \lor b) = a$;
  \item $0$ is the identity of $\lor$, and $1$ is the identity of $\land$;
  \item $\lor$ distributes over $\land$ and vice versa;
  \item $a \lor \neg a = 1$, and $a \land \neg a = 0$.
  \end{itemize}
\end{definition}

For clarity and succinctness, we will occasionally use three other operations
that can be defined using the original three\footnote{We use $+$ to denote
  symmetric difference because it is the additive operation of a Boolean ring.}:
\begin{align*}
  a \to b &= \neg a \lor b, \\
  a \leftrightarrow b &= (a \to b) \land (b \to a), \\
  a + b &= (a \land \neg b) \lor (\neg a \land b).
\end{align*}
We can also define a partial order $\le$ on $\mathbf{B}$ as $a \le b$ if $a = b
\land a$ (or, equivalently, $a \lor b = b$) for $a, b \in \mathbf{B}$.
Furthermore, let $a < b$ denote $a \le b$ and $a \ne b$. For the rest of this
paper, let $\mathbf{B}$ refer to the BA $(\mathbf{B}, \land, \lor, \neg, 0, 1)$.
For any $S \subseteq \mathbf{B}$, we write $\bigvee S$ for $\bigvee_{x \in S} x$
and call it the \emph{supremum} of $S$. Similarly, $\bigwedge S = \bigwedge_{x
  \in S} x$ is the \emph{infimum}. By convention, $\bigwedge \emptyset = 1$ and
$\bigvee \emptyset = 0$.

@@

\begin{lemma}[\cite{givant2008introduction}] \label{thm:representation}
  The following are equivalent:
  \begin{itemize}
  \item $\mathbf{B}$ is atomic.
  \item For any $x \in \mathbf{B}$,
    \[
      x = \bigvee_{\text{atoms } a \le x} a.
    \]
  \item $1$ is the supremum of all atoms.
  \end{itemize}
\end{lemma}

@@

\begin{definition}[\cite{gaifman1964concerning,DBLP:books/daglib/0090259}] \label{def:measure}
  A \emph{measure} on $\mathbf{B}$ is a function $m\colon
  \mathbf{B} \to \mathbb{R}_{\ge 0}$ such that:
  \begin{itemize}
  \item $m(0) = 0$;
  \item $m(x \lor y) = m(x) + m(y)$ for all $x, y \in \mathbf{B}$ whenever $x
    \land y = 0$.
  \end{itemize}
  If $m(1) = 1$, we call $m$ a \emph{probability measure}. Also, if $m(x) > 0$
  for all $x \ne 0$, then $m$ is \emph{strictly positive}.
\end{definition}

\begin{definition} \label{def:algebra_from_logic}
  Let $\mathcal{L}$ be a propositional (or first-order) logic, and let
  $\Delta$ be a theory in $\mathcal{L}$. We can define an equivalence
  relation on formulas in $\mathcal{L}$ as
  \[
    \alpha \sim \beta \quad \text{if and only if} \quad \Delta \vdash \alpha
    \leftrightarrow \beta
  \]
  for all $\alpha, \beta \in \mathcal{L}$. Let $[\alpha]$ denote the equivalence
  class of $\alpha \in \mathcal{L}$ with respect to $\sim$. We can then let
  $B(\Delta) = \{ [\alpha] \mid \alpha \in \mathcal{L} \}$ and define the
  structure of a BA on $B(\Delta)$ as
  \begin{align*}
    [\alpha] \lor [\beta] &= [\alpha \lor \beta], \\
    [\alpha] \land [\beta] &= [\alpha \land \beta], \\
    \neg[\alpha] &= [\neg\alpha], \\
    1 &= [\alpha \to \alpha], \\
    0 &= [\alpha \land \neg\alpha]
  \end{align*}
  for all $\alpha, \beta \in \mathcal{L}$. Then $B(\Delta)$ is the
  \emph{Lindenbaum-Tarski algebra} of $\Delta$
  \cite{koppelberg1989handbook,tarski1983logic}.
\end{definition}

\begin{figure}
  \[
    \begin{tikzcd}
      & & & & & \colorbox{color3}{$1$} \ar[dlll,dash,gray] \ar[dl,dash,gray]
      \ar[dr,dash,gray] \ar[drrr,dash,gray] & & & \\
      & & \colorbox{color3}{$p \lor q$} & & \colorbox{color3}{$q \to p$} & &
      \colorbox{color2}{$p \to q$} & & \colorbox{color4}{$\neg p \lor \neg q$}
      \\
      \colorbox{color3}{$p$} \ar[urr,dash,gray] \ar[urrrr,dash,gray] & &
      \colorbox{color2}{$q$} \ar[u,dash,gray] \ar[urrrr,dash,gray] & &
      \colorbox{color2}{$p \leftrightarrow q$} \ar[u,dash,gray]
      \ar[urr,dash,gray] & & \colorbox{color4}{$p + q$} \ar[ullll,dash,gray]
      \ar[urr,dash,gray] & & \colorbox{color4}{$\neg q$} \ar[ullll,dash,gray]
      \ar[u,dash,gray] & & \colorbox{color1}{$\neg p$} \ar[ullll,dash,gray]
      \ar[ull,dash,gray] \\
      & & \fcolorbox{black}{color2}{$p \land q$} \ar[ull,dash,gray]
      \ar[u,dash,gray] \ar[urr,dash,gray] & & \fcolorbox{black}{color4}{$p \land
        \neg q$} \ar[ullll,dash,gray] \ar[urr,dash,gray] \ar[urrrr,dash,gray] &
      & \fcolorbox{black}{color1}{$\neg p \land q$} \ar[ullll,dash,gray]
      \ar[u,dash,gray] \ar[urrrr,dash,gray] & & \fcolorbox{black}{color1}{$\neg
        p \land \neg q$} \ar[ullll,dash,gray] \ar[u,dash,gray]
      \ar[urr,dash,gray] \\
      & & & & & \colorbox{color1}{$0$} \ar[ulll,dash,gray] \ar[ul,dash,gray]
      \ar[ur,dash,gray] \ar[urrr,dash,gray] & & &
    \end{tikzcd}
  \]
  \[
    \begin{tikzcd}
      & \colorbox{color3}{$\left[1\right]$} \ar[dl,dash,gray] \ar[dr,dash,gray]
      & \\
      \fcolorbox{black}{color2}{$\left[q\right]$} & &
      \fcolorbox{black}{color4}{$\left[\neg q\right]$} \\
      & \colorbox{color1}{$\left[0\right]$} \ar[ul,dash,gray] \ar[ur,dash,gray]
      &
    \end{tikzcd}
  \]
  \caption{Two BAs from \cref{example:construction}: $B(\mathcal{L})$ at the top
    and $B(\Delta)$ at the bottom. An edge between elements $a$ and $b$ (with
    $a$ positioned lower than $b$) means that $a \le b$. Each element of
    $B(\Delta)$ is an equivalence class of elements of $B(\mathcal{L})$, and the
    colours show which elements of $B(\mathcal{L})$ belong to which class. In
    both algebras, atoms have borders around them.}
  \label{fig:example}
\end{figure}

\begin{example} \label{example:construction}
  Let $\mathcal{L}$ be a propositional logic with $p$ and $q$ as its only atoms.
  Then $L = \{ p, q, \neg p, \neg q \}$ is its set of literals. Let $w : L \to
  \mathbb{R}_{\ge 0}$ be the \emph{weight function} defined by
  \begin{align*}
    w(p) = 0.3, \\
    w(\neg p) = 0.7, \\
    w(q) = 0.2, \\
    w(\neg q) = 0.8.
  \end{align*}
  Let $\Delta$ be a theory in $\mathcal{L}$ with a sole axiom $p$. Then
  $\Delta$ has two models, i.e., $\{ p, q \}$ and $\{ p, \neg q \}$. The
  \emph{weighted model count} (WMC) \cite{DBLP:journals/ai/ChaviraD08} of $\Delta$ is
  then
  \[
    \sum_{\omega \models \Delta} \prod_{\omega \models l} w(l) =
    w(p)w(q) + w(p)w(\neg q) = 0.3.
  \]

  The corresponding BA $B(\Delta)$ can then be constructed using
  \cref{def:algebra_from_logic}. Alternatively, one can first construct the free
  BA generated by the set $\{ p, q \}$---this corresponds to $B(\mathcal{L})$ in
  \cref{fig:example}---and then take a quotient with respect to either the
  filter generated by $p$ or the ideal\footnote{More details on these concepts
    can be found in many books on BAs
    \cite{givant2008introduction,koppelberg1989handbook}.} generated by $\neg
  p$. In any case, the resulting BA is pictured at the bottom of
  \cref{fig:example}. To summarise, 
\end{example}
% TODO: could talk about how each element can be seen as a set of models.
% TODO: the number of elements in the algebra.

% TODO: Feedback: Probably a good idea to explain how elements of the Boolean
% algebra map to models in WMC formulations. Probably do an example:
% 0.3 p; 0.2 q => M = <p, qbar> has weight 0.3 * (1-0.2)
% then it corresponds to an atom in an algebra defined over the language defined
% over propositions p and q.
% TODO: connect the two definitions of WMC

\begin{definition} \label{def:wmc}
  Let $\mathbf{B}$ be an atomic BA, and let $M \subset \mathbf{B}$ be its set of
  atoms. Let $L \subset \mathbf{B}$ be such that every atom $m \in M$ can be
  uniquely expressed as $m = \bigwedge L'$ for some $L' \subseteq L$, and let
  $w\colon L \to \mathbb{R}_{\ge 0}$ be arbitrary. The \emph{weighted model
    count} $\WMC_w\colon \mathbf{B} \to \mathbb{R}_{\ge 0}$ is defined as
  \[
    \WMC_w(x) = \begin{cases}
      0 & \text{if } x = 0 \\
      \prod_{l \in L'} w(l) & \text{if } M \ni x = \bigwedge L' \\
      \sum_{\text{atoms } a \le x} \WMC_w(a) & \text{otherwise}
    \end{cases}
  \]
  for any $x \in \mathbf{B}$. Furthermore, we define the \emph{normalised
    weighted model count} $\nWMC_w\colon \mathbf{B} \to [0, 1]$ as $\nWMC_w(x) =
  \frac{\WMC_w(x)}{\WMC_w(1)}$ for all $x \in \mathbf{B}$. For both $\WMC_w$ and
  $\nWMC_w$, we will drop the subscript when doing so results in no potential
  confusion.
\end{definition}

% TODO: Feedback: clarify the relationship between NWMC and the probability
% induced by a WMC formulation, i.e., Pr_{\Delta, w}(q) = \frac{WMC(\Delta \land
% q, w)}{WMC(\Delta, w)}.

\begin{proposition}
  $\WMC$ is a measure and $\nWMC$ is a probability measure.
\end{proposition}
\begin{proof}
  First, note that $\WMC$ is non-negative and $\WMC(0) = 0$ by definition. Next,
  let $x, y \in \mathbf{B}$ be such that $x \land y = 0$. We want to show that
  \begin{equation} \label{eq:additivity_proof}
    \WMC(x \lor y) = \WMC(x) + \WMC(y).
  \end{equation}
  If, say, $x = 0$, then \cref{eq:additivity_proof} becomes
  \[
    \WMC(y) = \WMC(0) + \WMC(y) = \WMC(y)
  \]
  (and likewise for $y = 0$). Thus we can assume that $x \ne 0 \ne y$ and use
  \cref{thm:representation} to write
  \[
    x = \bigvee_{i \in I} x_i \quad \text{and} \quad y = \bigvee_{j \in J} y_j
  \]
  for some sequences of atoms $(x_i)_{i \in I}$ and $(y_j)_{j \in J}$. If
  $x_{i'} = y_{j'}$ for some $i' \in I$ and $j' \in J$, then
  \[
    x \land y = \bigvee_{i \in I} \bigvee_{j \in J} x_i \land y_j = x_{i'} \land
    y_{j'} \ne 0,
  \]
  contradicting the assumption. This is enough to show that
  \begin{align*}
    \WMC(x \lor y) &= \WMC\left( \left( \bigvee_{i \in I} x_i \right) \lor \left(\bigvee_{j \in J} y_j \right) \right) = \sum_{i \in I} \WMC(x_i) + \sum_{j \in J} \WMC(y_j) \\
                   &= \WMC(x) + \WMC(y),
  \end{align*}
  finishing the proof that $\WMC$ is a measure. This immediately shows that
  $\nWMC$ is a probability measure since, by definition, $\nWMC(1) = 1$.
\end{proof}

@@

\begin{proposition}
  Let $\mathbf{B}$ be a finite measure algebra with measure $m\colon \mathbf{B} \to
  \mathbb{R}_{\ge 0}$. Let $L \subset \mathbf{B}$ be defined as
  \[
  L = \{ l_i \mid i \in [n] \} \cup \{ \neg l_i \mid i \in [n] \}
  \]
  for some $n \in \mathbb{N}$. Finally, assume that $\mathbf{B}$ has $2^n$
  atoms, where each atom $a \in \mathbf{B}$ is an infimum
  \[
    a = \bigwedge_{i=1}^n a_i
  \]
  such that $a_i \in \{ l_i, \neg l_i \}$ for $i \in [n]$. Then there exists a
  weight function $w\colon L \to \mathbb{R}_{\ge 0}$ that makes $m$ a WMC measure if
  and only if
  \begin{equation} \label{eq:wmccondition}
  m(l \land l') = m(l)m(l')
  \end{equation}
  for all distinct $l, l' \in L$ such that $l \ne \neg l'$.
\end{proposition}

@@

\begin{proof}
  Let us begin with the `if' part of the statement. Let $w\colon L \to
  \mathbb{R}_{\ge 0}$ be defined by
  \begin{equation} \label{eq:assumption}
    w(l) = m(l)
  \end{equation}
  for all $l \in L$. We are going
  to show that $\nWMC = m$. First, note that $\nWMC(0) = 0 = m(0)$ by the
  definitions of both $\nWMC$ and $m$. Second, let
  \begin{equation} \label{eq:def_of_a}
    a = \bigwedge_{i=1}^n a_i
  \end{equation}
  be an atom in $\mathbf{B}$ such that $a_i \in \{ l_i, \neg l_i \}$ for all $i
  \in [n]$. Then
  \[
    \nWMC(a) = \frac{\WMC(a)}{\WMC(1)} = \frac{1}{\WMC(1)} \prod_{i=1}^n w(a_i)
    = \frac{1}{\WMC(1)} \prod_{i=1}^n m(a_i) = \frac{1}{\WMC(1)} m \left(
      \bigwedge_{i=1}^n a_i \right) = \frac{m(a)}{\WMC(1)}
  \]
  by \cref{def:wmc,eq:assumption,eq:wmccondition,eq:def_of_a}. Now we just need
  to show that $\WMC(1) = 1$. Indeed,
  \begin{align*}
    \WMC(1) &= \sum_{\text{atoms } a \in \mathbf{B}} \WMC(a) = \sum_{\text{atoms
      } a \in \mathbf{B}} \prod_{i=1}^n w(a_i) = \sum_{\text{atoms } a \in
      \mathbf{B}} \prod_{i=1}^n m(a_i) \\
    &= \sum_{\text{atoms } a \in
      \mathbf{B}} m \left( \bigwedge_{i=1}^n a_i \right) = \sum_{\text{atoms } a
      \in \mathbf{B}} m(a) = m \left( \bigvee_{\text{atoms } a \in \mathbf{B}}
    \right) = m(1) = 1.
  \end{align*}
  Finally, note that if $\nWMC$ and $m$ agree on all atoms, then they must also
  agree on all other non-zero elements of the Boolean algebra.

  For the other direction, we are given a weight function $w\colon L \to
  \mathbb{R}_{\ge 0}$ that induces a measure $m = \nWMC\colon \mathbf{B} \to
  \mathbb{R}_{\ge 0}$, and we want to show that \cref{eq:wmccondition} is
  satisfied. Let $k_i, k_j \in L$ be such that $k_i \in \{ l_i, \neg l_i \}$,
  $k_j \in \{ l_j, \neg l_j \}$, and $i \ne j$. We will first prove an auxiliary
  result that
  \begin{equation} \label{eq:to_prove}
    m(k_i \land k_j) = m(k_i)m(k_j)
  \end{equation}
  is equivalent to
  \begin{equation} \label{eq:to_prove2}
    m(k_i \land k_j) \cdot m(\neg k_i \land \neg k_j) = m(k_i \land \neg k_j)
    \cdot m(\neg k_i \land k_j).
  \end{equation}
  First, note that $k_i$ can be expressed as
  \[
    k_i = (k_i \land k_j) \lor (k_i \land \neg k_j)
  \]
  with the condition that
  \[
    (k_i \land k_j) \land (k_i \land \neg k_j) = 0,
  \]
  so, by properties of a measure,
  \begin{equation} \label{eq:temp}
    m(k_i) = m(k_i \land k_j) + m(k_i \land \neg k_j).
  \end{equation}
  Applying \cref{eq:temp} and the equivalent expression for $m(k_j)$ allows us
  to rewrite \cref{eq:to_prove} as
  \begin{align*}
    m(k_i \land k_j) &= [m(k_i \land k_j) + m(k_i \land \neg k_j)] \cdot [m(k_i \land k_j) + m(\neg k_i \land k_j)] \\
                     &= m(k_i \land k_j)^2 + m(k_i \land k_j)[m(k_i \land \neg k_j) + m(\neg k_i \land k_j)] + m(k_i \land \neg k_j)m(\neg k_i \land k_j)
  \end{align*}
  Dividing both sides by $m(k_i \land k_j)$ gives
  \begin{equation} \label{eq:temp2}
    1 = m(k_i \land k_j) + m(k_i \land \neg k_j) + m(\neg k_i \land k_j) +
    \frac{m(k_i \land \neg k_j)m(\neg k_i \land k_j)}{m(k_i \land k_j)}.
  \end{equation}
  Since $k_i \land k_j \land k_i \land \neg k_j = 0$, and
  \[
    (k_i \land k_j) \lor (k_i \land \neg k_j) = k_i \land (k_j \lor \neg k_j) =
    k_i \land 1 = k_i,
  \]
  we have that
  \[
    m(k_i \land k_j) + m(k_i \land \neg k_j) = m(k_i).
  \]
  Similarly, $k_i \land \neg k_i \land k_j = 0$, and
  \[
    k_i \lor (\neg k_i \land k_j) = (k_i \lor \neg k_i) \land (k_i \lor k_j) =
    k_i \lor k_j,
  \]
  so
  \[
    m(k_i) + m(\neg k_i \land k_j) = m(k_i \lor k_j).
  \]
  Finally, note that
  \[
    (k_i \lor k_j) \land \neg(k_i \lor k_j) = 0,
  \]
  and
  \[
    (k_i \lor k_j) \lor \neg(k_i \lor k_j) = 1,
  \]
  so
  \[
    m(k_i \lor k_j) + m(\neg(k_i \lor k_j)) = m(1) = 1.
  \]
  This allows us to rewrite \cref{eq:temp2} as
  \[
    \frac{m(k_i \land \neg k_j)m(\neg k_i \land k_j)}{m(k_i \land k_j)} = 1 -
    m(k_i \lor k_j) = m(\neg(k_i \lor k_j)) = m(\neg k_i \land \neg k_j)
  \]
  which immediately gives us \cref{eq:to_prove2}.

@@

A well-known way to overcome this limitation of independence is by adding more
literals \cite{DBLP:journals/ai/ChaviraD08}, i.e., extending the set $L$ covered
by the WMC weight function $w\colon L \to \mathbb{R}_{\ge 0}$. Let us translate this
idea to the language of Boolean algebras.
